SHARP - II: Revealing a bias in observational measurements of dark matter substructure with gravitational lens flux ratios

### 2015-12-06 papers/B1555_EdgeOn/B1555_letter.tex
Updates to the paper text and plotting code

The next step is to test if a physically plausible edge-on disc can cause a perturbation in the strong lensing mass model that produces the observed flux ratio anomaly of  merging images A and B. An exponential disc profile best describes the disc component in spiral galaxies because it matches the light distribution \citep{Kee98}. We thus choose a SIE plus an exponential disc model (SIE+ExpDisc) in our study of B1555+375. The free parameters of the model are the SIE Einstein radius ($b$), centroid position, ellipticity ($e$), position angle ($\theta$), and the exponential disc intrinsic central density ($\kappa_0$) and scale length ($R_s$). A detailed definition of these mass profiles is described by \citet{Kee01}. The source position adds two additional model parameters. {\bf We have not included a third component such as an external shear or additional NFW halo term because the number of constraints provided by the compact images detected by MERLIN is not sufficient to constrain such a model.} 
%An SIE profile is also sufficient to include the halo mass since B1555 has a small Einstein radius.} 
Significantly deeper observations at a higher bandwidth with current very long baseline interferometric arrays can test more complex models in the future by imaging the extended gravitational arc and more significantly detecting image D.

@@

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:model}. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in fluxes may due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the ExpDisc component makes up about 10 per cent of the total {\bf projected} mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$).  {\bf In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total enclosed mass \citep[e.g.][]{swells2}.  In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the convergence of the ExpDisc component as a good description of the baryonic mass of the lens galaxy.}  We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models need the full extent of the gravitational arc for constraints. {\bf The extinction in the AO imaging is so severe that components B and D are not visible, and the arc that connects A and B appears to be truncated.  The low signal-to-noise ratio and the strong extinction in lens galaxy prevent us from using modeling techniques that incorporate the extended NIR emission \citep[e.g.][]{V12}.  Future high resolution radio imaging is needed in order to place stronger constraints on the mass model.}

@@

\textbf{ Our analysis of B1555+375 demonstrates that the choice of the macromodel used to describe the `smooth' lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to a misinterpretation of the origin of flux-ratio anomalies in some systems. Before concluding that the flux-ratio anomalies in any given lens system are the result of dark matter substructure, therefore, it is critical that other possible sources of significant curvature in the lensing potential (i.e., `sharp' mass features such as edge-on disks or even massive luminous satellites) are taken into account 
\citep[e.g.][]{Xu15}. The SWELLS survey, for example, presents a sample of 16 edge-on disks from SDSS \citep{swells1}.  This result is timely, as upcoming large-scale surveys (e.g., the Dark Energy Survey, Kilo-Degree Survey, Euclid, and the Large Synoptic Survey Telescope) will deliver orders of magnitude more gravitationally lensed quasars for which flux ratios can
  be measured using narrow lines \citep[e.g.][]{N14}.}
  %Finally, we note that massive edge-on discs are \textit{not} observed in some flux-ratio anomaly systems \citep[and the presence of a disc-like component may not resolve the anomaly anyway;][]{evans03}, indicating that these systems are still useful for probing the dark matter substructure mass function.

>>>>

The next step is to test if a physically plausible edge-on disc can cause a perturbation in the strong lensing mass model that produces the observed flux ratio anomaly of  merging images A and B. An exponential disc profile best describes the disc component in spiral galaxies because it matches the light distribution \citep{Kee98}. We thus choose a SIE plus an exponential disc model (SIE+ExpDisc) in our study of B1555+375. The free parameters of the model are the SIE Einstein radius ($b$), centroid position, ellipticity ($e = 1 - q$), position angle ($\theta$), and the exponential disc intrinsic central density ($\kappa_0$) and scale length ($R_s$). A detailed definition of these mass profiles is described by \citet{Kee01}. The source position adds two additional model parameters. {\bf We have not included a third component such as an external shear or additional NFW halo term because the number of constraints provided by the compact images detected by MERLIN is not sufficient to constrain such a model.} 
%An SIE profile is also sufficient to include the halo mass since B1555 has a small Einstein radius.} 
{\bf 
More sophisticated mass models require additional
constraints, such as those from the lensed extended emission in either
the NIR or radio imaging.  While we have successfully used the former
approach in the past \citep[e.g.][]{V12}, for B1555+375 the extinction
in the AO imaging is so severe that components B and D are not
visible, and the arc that connects A and B appears to be truncated.
The low signal-to-noise ratio in the ring away from the lensed AGN and
the strong extinction in lens galaxy prevent us from using this
approach.  What are, thus, needed are significantly deeper
observations at a higher bandwidth with very long baseline
interferometric arrays, which can test more complex models in the
future by imaging the extended gravitational arc and more
significantly detecting image D. }

@@

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:model}. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in fluxes may due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the ExpDisc component makes up about 10 per cent of the total {\bf projected} mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$).  {\bf In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total enclosed mass \citep[e.g.][]{swells2}.  In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy.}  We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require {\bf the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.}

@@

\textbf{ Our analysis of B1555+375 demonstrates that the choice of the macromodel used to describe the `smooth' lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to a misinterpretation of the origin of flux-ratio anomalies in some systems. Before concluding that the flux-ratio anomalies in any given lens system are the result of dark matter substructure, therefore, it is critical that other possible sources of significant curvature in the lensing potential (i.e., `sharp' mass features such as edge-on disks or even massive luminous satellites) are taken into account 
\citep[e.g.][]{Xu15}. A number of systems with significant disc components are
already known.  These include, for example, the sample of 16 edge-on disks from SDSS found in the SWELLS survey \citep{swells1}.  This result is timely, as upcoming large-scale surveys (e.g., the Dark Energy Survey, Kilo-Degree Survey, Euclid, and the Large Synoptic Survey Telescope) will deliver orders of magnitude more gravitationally lensed quasars for which flux ratios can
  be measured using narrow lines \citep[e.g.][]{N14}.}
  %Finally, we note that massive edge-on discs are \textit{not} observed in some flux-ratio anomaly systems \citep[and the presence of a disc-like component may not resolve the anomaly anyway;][]{evans03}, indicating that these systems are still useful for probing the dark matter substructure mass function.
